---
title: "StackOverFlow 2023 Analysis"
author: "Hesam Korki"
date: "22 Nov. 2023"
output: html_document
---

> There is a renv.lock file provided in the repo, we'd install the required packages and then laod the library

```{r}
library("tidyverse")
library("eurostat")
library("countrycode")
```
# Loading Data

```{r}
data <- read_csv("data/survey_results_public.csv", show_col_types = FALSE)
eu_wage <- read_tsv("data/earn_mw_cur_tabular.tsv", show_col_types=FALSE)
```

## Configuring Plotting

```{r}
# defining a theme for adjusting spacings of graphs
adjust_spacing <-
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 20)),
    plot.margin = margin(t=20, r=50, b=20, l=20, unit = "pt")
  )

primary_color <- "#429696"
```


# Exploration

:::{.callout-note appearance="minimal"}
There are $89,184$ rows and $84$ columns

- Each row represents the answers of a participant
- Each column corresponds to a question of the survey 
:::

```{r}
glimpse(data)
```

:::{.callout-warning appearance="minimal"}
Most columns have values as a single string which contains semicolon separated values which is not as tidy as we would want.
:::


What interesting insights do you get by looking at this graph?


```{r}
data |>
  ggplot() +
  aes(y = fct_rev(fct_infreq(Age))) +
  geom_bar(fill = primary_color, alpha = 0.8) +
  geom_text(
    aes(label = scales::percent(after_stat(count)/sum(count))),
    stat = "count",
    hjust = -0.2,
    size = 3,
    color = "black"
  ) +
  labs(x = NULL, y = NULL, title = "How old are the participants?") +
  annotate(
    "label",
    x = 32000, y = 1.5,
    label = paste("Total Number of Participents:", scales::comma(nrow(data))),
    color = "black",
    size=3
  ) +
  lims(x = c(0, 40000)) +
  theme_minimal() +
  adjust_spacing + 
  scale_x_continuous(labels = scales::comma)
  
```

```{r}

employment_df <- data |> 
  select(ResponseId, Employment) |> 
  separate_rows(Employment, sep=";") |> 
  mutate(Employment=trimws(Employment)) |> 
  drop_na(Employment)


employment_df |>
  ggplot() +
  aes(y = fct_rev(fct_infreq(Employment))) +
  geom_bar(fill = primary_color, alpha = 0.8) +
  geom_text(
    aes(label = scales::percent(after_stat(count)/sum(count))),
    stat = "count",
    hjust = -0.2,
    size = 3,
    color = "black"
  ) +
  labs(x = NULL, y = NULL, title = "Employment status of respondants") +
  lims(x = c(0, 70000)) +
  theme_minimal() +
  adjust_spacing + 
  scale_x_continuous(labels = scales::comma)
```
```{r}

continent_data <- data |>
  mutate(Continent = countrycode(
    sourcevar = Country,
    origin = "country.name",
    destination = "continent"
  )) |>
  drop_na(Continent) |> 
  group_by(Continent, Country) |> 
  summarise(freq= n()) |> 
  mutate(
    Country = case_when(
      Country == "United States of America" ~ "USA",
      Country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK",
      TRUE ~ Country
    )
  )

continent_data |>
  ggplot() +
  aes(y = Continent,
      x = freq,
      fill = fct_lump(Country, n=10, w=freq)) +
  geom_col() +
  labs(title = "Diversity of Respondants", x = NULL, y = NULL, fill=NULL) +
  theme_minimal()+ 
  scale_fill_brewer(palette = "Set3") +
  adjust_spacing +
  scale_x_continuous(labels = scales::comma)

```
```{r}
continent_data |> 
  arrange(desc(freq))
  
```





```{r}
eu <- data |> 
  filter(Country %in% eu_countries$name)


eu |>
  ggplot() +
  aes(y = fct_rev(fct_infreq(Country))) +
  geom_bar(fill = primary_color, alpha = 0.8) +
  geom_text(
    aes(label = scales::percent(after_stat(count)/sum(count))),
    stat = "count",
    hjust = -0.2,
    size = 3,
    color = "black"
  ) +
  labs(x = NULL, y = NULL, title = "How old are the participants?") +
  annotate(
    "label",
    x = 8000, y = 3,
    label = paste("Total Number of Participents:", scales::comma(nrow(eu))),
    color = "black",
    size=3
  ) +
  lims(x=c(0,10000))+
  theme_minimal() +
  adjust_spacing
```

